A Global Genome-wide Scan with Optimal Cutoff Mining for Emerging Biomarkers in Head and Neck Squamous Cell Carcinoma

### 2021-01-02 main.tex
\maxdeadcycles = 300

%=================================================================
\documentclass[cancers,article,submit,moreauthors,pdftex]{Definitions/mdpi}

@@

\subsection*{Patient Cohort}

@@

\subsubsection*{RNA Sequencing Data}

@@

\subsubsection*{Clinical Data}

@@

\subsection*{Cutoff Finder Core Engine}

\subsection*{Statistical Consideration}

\subsection*{Biomarker Selection}

>>>>

%=================================================================
\documentclass[cancers,article,submit,moreauthors,pdftex]{Definitions/mdpi}
\maxdeadcycles=300 % Output loop---200 consecutive dead cycles.

@@

\subsection{Patient Cohort}

@@

\subsubsection{RNA Sequencing Data}

@@

\subsubsection{Clinical Data}

@@

\subsection{Cutoff Finder Core Engine}

%\label{key:cutoff}
To evaluate the effect of gene expression on the patient's survival, we introduced the stratifying of patients with Kaplan-Meier survival analysis according to each gene's low/high expression.
Our cutofFinder\_func subroutine employs the minimum P-value approach to recognizing cutoff points in continuous gene expression measurement for patients sub-population.
First, patients were ordered by \acrshort{rnaseq} value (RSEM) of a given gene. Next, patients were stratified at a serial cut (counted by person ranked in 30\% to 70\% percentile of the cohort; please see Figure \ref{fig:figure1} "HNSCC cohort"). The survival risk differences of the two groups were estimated by log-rank test to yield around 165 Kaplan-Meier P-values for each gene.
Then, the optimal cutoff of \acrshort{rnaseq}, giving the minimum P-value, was selected by the cutofFinder\_func subroutine.
This iteration method could calculate all possible cutoff of each gene expression in this cohort. At each run of cutofFinder\_func function call for an individual gene, it returned an optimal cutoff value (e.x. 0.027 for gene \acrlong{CAMK2N1}, \acrshort{CAMK2N1}). The optimal cutoff value and its correlated patient grouping size (e.x. low-expression in 262 persons vs. high-expression in 152 persons with gene \acrshort{CAMK2N1}) would be returned to the main program to allow downstream Cox survival analysis. The percentile range we applied as 30\% to 70\% was used to avoid a small grouping effect\cite{Miller1982}\cite{Mizuno2009a}.
In case there was no significant P-value, a median expression of this gene was set as its cutpoint as usual.

\subsection{Statistical Consideration}

Our workflow has loops to call function survival\_marginSFP(GeneName) with given GeneName to process the survival analysis gene by gene.
We dichotomized the clinicopathological features, which includes gender (male/ female), age at diagnosed (below/beyond 65 year-old), clinical tumor size (T1-2/T3-4), clinical nodal status (negative/positive), clinical distant metastasis (negative/positive), TNM staging (early/late), surgical margin status (negative/positive) and tobacco exposure (low/high). The patients were grouped by an \acrshort{rnaseq} value of each gene, cut at low- or high-expression on an optimal P-value obtained from the cutofFinder\_func subroutine (see the section of "Cutoff Finder Core Engine"). Pearson's chi-square test was used for these binary variables. Kaplan-Meier survival was analyzed using the log-rank test for two groups OS comparison.
The Cox proportional hazards regression is the widely accepted approach for modeling survival while accounting for confounding factors\cite{Magen2019}. Univariate and multivariate Cox proportional regression model\cite{Andersen1982}, using the "coxph" function in R package "survival", was applied to calculate hazard ratio, \acrfull{ci95} and its significance, and to estimate the independent contributions of each clinicopathological features to the OS.
Results were considered statistically significant when a two-sided P-value $<$ 0.05, or a lower threshold if indicated.
The \acrfull{fdr} ($< 0.05$) could be used to pick up the optimal P-value to ensure the control for type I error of the Kaplan-Meier survival test during the cutoff finding procedure.
%statistical inference problems in biomedical and genomic data analysis routinely involve the simultaneous test of thousands, or even millions, of null hypotheses. Examples
There were also multiple correlated tests of null hypotheses in the Kaplan-Meier family during our global scanning of protein-coding genes. The stringent Bonferroni correction could result in an adjusted P-value to ensure the control for type I error. 
%There were multiple correlated tests in the family of Kaplan-Meier survival hypotheses during our global scanning of protein-coding genes. The stringent Bonferroni correction could result in an adjusted P-value to ensure the control for type I error.

The resulting data, including Kaplan-Meier curves, cumulative P-value plots, and Cox regression tables, were exported to ".xlsx" and ".Rda" files (by R package "r2excel") for subsequent biomarker selection.

\subsection{Biomarker Selection}
